A Test of the Frequency Independence Assumption of the Power System Parameters used in Geomagnetically Induced Current Calculations

### 2019-02-26 paper.tex
Updates from Overleaf

\begin{abstract}
A common assumption used in estimating geomagnetically induced currents (GICs) in a power system given a nearby direct measurement or indirect estimate (via a transfer function with input of the geomagnetic field) of the horizontal geoelectric field components $E_x$ and $E_y$ on Earth's surface is that the system is resistive. That is, the approximation $GIC(t)$  $=$ $a_oE_x(t) + b_oE_y(t)$ (Model 1) is used, where $a_o$ and $b_o$ are frequency independent network parameters.  We provide the first test of this assumption using GIC measurements made in a 187 kV transformer connected to a $\sim$100~km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$10~km \todo[color=blue!40]{Shinichi: Is 10 km correct?} away.  We compute frequency dependent network parameters $a(\omega)$ and $b(\omega)$ in Model 2, $GIC(\omega)$ $=$ $a(\omega)E_x(\omega) + b(\omega)E_y(\omega)$, and find that they are frequency dependent and that this model provides a significantly better representation of measured GIC.
\end{abstract}

@@

Geomagnetically induced currents (GICs) are electric currents in conducting systems due to electric fields induced near Earth's surface. These fields are a result of time variations in ionospheric currents on timescales on the order of hours \citep{Ohtani2000} and the movement of Earth's surface relative to slow-varying current systems in Earth's ionosphere that are near stationary relative to the Sun \citep{Stening2013}. Of particular interest are currents induced in electric power systems as they can lead to system degradation, disruption, and failure \citep{Albertson1993,NERC2012}. Accurate estimation of GICs given direct measurements of the electric field (or estimates of the electric field based on the geomagnetic field) is important for power system design, retrospective analysis, and mitigation of the impacts of space weather on power systems \citep{Molinski2002,Thomson2010,NERC2012,Gaunt2014}. \todo{Pierre and Trevor - any other reference suggestions?}

@@

\cite{Oyedokun2013a} developed a method to account for the finite inductive response time of transformers by taking the GIC predicted based on the quasi-DC assumption and then applying a piecewise-linear L/R filter derived from bench-scale transformer measurements, with a time constant dependent on load, to obtain an alternative estimate of GIC.

\todo[inline]{What can I conclude from \cite{Oyedokun2013a}? Mostly the thesis documents visual differences in time series traces with and without accounting for transformers.}

In this work, we use a unique dataset in which measurements $\mathbf{E}$ and the surface horizontal geomagnetic field, $\mathbf{B}$, were made near a site where GICs were being measured in an electric power system.  We first estimate the system parameters $a_o$ and $b_o$ using conventional methods, which assume that they are frequency independent.  Then, we use a model in which the system parameters are frequency dependent, compute the frequency dependence, and compare this model with the traditional frequency independent model.

@@

The only modification made to the magnetic field measurements was the replacement of 36 data points in all components of the magnetic field on 2006-08-06 with linearly interpolated values. The values for $B_x$ and $B_y$ were the same and much larger than the surrounding values.

@@

\noindent
the matrix equation $\mathbf{GIC} = \underline{\mathbf{E}}\cdot\mathbf{p}$ is created, where $\mathbf{GIC}$ is a 864000x1 matrix of 1 day of observations, $\mathbf{p} = [a_o,b_o]^T$, and $\underline{\mathbf{E}}$ is a 86400x2 matrix with rows of $[E_x(t), E_y(t)]$ observations. The values of $\mathbf{p}$ are then obtained using a standard linear regression solver. The values of $a_o$ and $b_o$ determined in this way correspond to those that minimize the sum-of-squares of $GIC(t)-G_o(t)$ over a 1-day interval. \citep[][provided the mathematically equivalent closed-form equations.]{Pulkkinen2007}. The averages $\overline{a}_o$ and $\overline{b}_o$ are the average of the 35 $a_o$ and $b_o$ values computed using the 35 unique 1-day window segments of available data.

In Model 2, the system parameters $a$ and $b$ are frequency dependent according to

@@

The same approach is used for solving for $z_x$ and $z_y$ in Equation~\ref{model4}. For the transfer function $\boldsymbol{Z}$ in $\mathbf{E}(\omega) = \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$ used in Model~3, the $xx$ and $yx$ components of $\boldsymbol{Z}$ are solved for using $E_x(\omega) = Z_{xx}(\omega)B_x(\omega) + Z_{yx}(\omega)B_{y}(\omega)$; the $yx$ and $yy$ components are solved for using $E_y(\omega) = Z_{yx}(\omega)B_x(\omega) + Z_{yy}(\omega)B_y(\omega)$.

@@

There are many statistical methods and variations thereof that have been developed for estimating transfer functions in the magnetotelluric (MT) literature that use additional steps and processing \citep{Egbert1986,Chave1987,Chave1989,Jones1989,Larsen1996,Egbert1997,Eisel2001,Chave2004,Chave2012,Chave2017} beyond the ones used that follow the basic method in \cite{Sims1971,Simpson2005}. These objective of these works is typically the estimation of a robust and low bias transfer function. In the results presented, we did not use a robust statistical method or attempt to correct for bias in estimates as our primary interest is comparing the ability of different models to predict the data (in a mean-squared-error sense) rather than estimating the model parameters. As discussed and evaluated in~\cite{Weigel2017}, a model developed using robust and bias correction methods will not necessarily be an optimal prediction model (when the sum-of-squares error is used as a prediction metric).

@@

\begin{table}
\caption{Out-of-sample model performance metrics for models $m=1-4$. MSE$_m$/MSE$_1$ is the ratio of the MSE for model $m$ to that of Model 1. The PE and MSE ratios are the average of 35 values and the confidence interval, CI, was determined from 1000 bootstrap samples of these 35 values.}
\centering
\begin{tabular}{l c c c c}
\hline
$m$\hspace{1em} Model & PE & 95\% CI & MSE$_m$/MSE$_1$ & 95\% CI\\
\hline
1\hspace{1em} $G_o(t) = a_oE_x(t) + b_oE_y(t)$ & 0.35 & [0.25, 0.45] & 1.0 & \\
2\hspace{1em} $G_E(t) = A(\omega)E_x(t) + B(\omega)E_y(t)$ & 0.60 & [0.54, 0.65] & 1.8 & [1.6, 2.0]\\
3\hspace{1em} $G_{E'}(t) = A(\omega)E'_x(t) + B(\omega)E'_y(t)$ & 0.78 & [0.76, 0.80] & 3.1 & [2.8, 3.4]\\
4\hspace{1em} $G_{B}(t) = Z_xB_x(t) + Z_yB_y(t)$ & 0.83 & [0.80, 0.85] & 4.6 & [3.8, 5.3]\\
\hline
\end{tabular}
\end{table}

Figure~\ref{predictions} shows a 1-day interval for which the value of the prediction efficiency for each model was in the same order as that shown in Table~1, i.e., Model 1 has the lowest PE and Model 4 has the lowest.

The primary conclusions of this work follow from the results shown in Table~1: the frequency dependent model, Model 2, is significantly better at predicting GIC than the frequency independent model, Model 1. Model 3, which uses horizontal magnetic field measurements as its input provides better estimates of GIC than Model 2. Model 4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides the best predictions of GIC.

@@

Consistent with the results in Table~1, The SN ratio for Model 1 is lower than that for Models 2-4 at all frequencies.  Above $10,000$~s (2.8~hrs), the SN ratio ordering is the same as the ordering of the PE and MSE ratios shown in Table~1, with Model~1 having the lowest and Model~4 having the highest SN ratio. Below $500$~s (8~min), Model 2 has the largest SN ratio. This result is somewhat visible in the sample interval shown in Figure~\ref{predictions} where the high frequency fluctuations in the prediction error for Model 2 are smaller than that for Models 3 and 4.

The average parameters for Model 1 are $a_o = $ $71.9 \pm 27.8$ A/(V/km) and $b_o =$ $-70.0 \pm 65.8$ A/(V/km) and the distribution of the 35 parameters calculated using each 1-day interval is shown in Figure~\ref{histogram}. (Results are similar if 1-minute averages are used; \cite{Watari2015} found \todo[color=blue!40]{Shinichi: I get $a_o=45.8$ and $b_o=-59.9$ in this time interval.} $a_o=38.1$ A/(V/km) and $b_o=-7.4$ A/(V/km) for the interval 2006-12-14--2006-12-15 JST time interval).

The wide distribution of parameter values and large error bars are suggestive that Model~\ref{model1} is a poor representation of the measurements. However, the model does have a good prediction performance - the PE is 0.35 and the data/model correlation coefficient of 0.64. In a similar way that $|\mathbf{B}|(t)$ or $d|\mathbf{B}|(t)/dt$ are often good predictors of $GIC(t)$ \todo{Pierre: Any suggestions?} {\color{red} \textbf{[REFS]}} despite the fact that physically $\mathbf{E}$ drives GIC and $\mathbf{E}$ is related to $\mathbf{B}$ through a frequency dependent transfer function, Model 1 is a fair predictor of GICs, but is a poor approximation relative to Models 2 and 3.

@@

\section{Discussion and Conclusions}

The primary implications of the results presented are (1) Significantly improved predictions of GICs can be made if frequency dependent system parameters are used and (2) A direct comparison of $a_o$ and $b_o$ estimated using Equation~\ref{model1} and computed using power system configuration information is problematic because the data are not consistent with the assumptions of Equation~\ref{model1}. Although the use of $a_o$ and $b_o$ values estimated from the data give a reasonable out-of-sample prediction model, it is not known how much these data estimated values will differ from those directly calculated.

@@

Ideally, frequency dependent system parameters could be calculated analytically so that GIC measurements are not needed for their estimates. Because the details of the power system under consideration in this work are not available, we were not able make this comparison. Note that frequency independent system parameters calculated using the approach in this work may not be directly comparable to those found from a network analysis for several reasons: (1) At periods less than $\sim 30$~s, the signal-to-noise ratio is low and the parameter estimates are expected to be biased; as shown in the Appendix, the transfer function estimates are highly sensitive to the method used for their calculation at (2) Also at periods less than $\sim 30$~s, effects of the instrument response may be significant. \cite{Fujii2015} has noted that the 1-s geoelectric field data from the Memambetsu Observatory were low-pass filtered with a 3.3-s cutoff-period prior to 2010; (3) The GIC observed in a power system is based on the integration of the electric field along a power line. All of the models assume that the electric field is constant along the line where the GIC is measured, but many works have shown that the electric field is expected to have variations on length scales comparable to the line considered ($\sim$100~km). \todo{REFS}

@@

\acknowledgments
Magnetic and electric field data made at the Kakioka Magnetic Observatory was obtained from the Japan Meteorological Agency data portal \url{http://www.kakioka-jma.go.jp/obsdata/metadata/en}. \todo[color=blue!40]{Shinichi: How should this be referenced?} GIC data provided by ...

%\clearpage

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_raw_All_20060805.pdf}
\caption{(Top) One-second-cadence geomagnetic field measured at the Memanbetsu Magnetic Observatory (MMB). (Middle) One-second-cadence geoelectric field measured at the Memanbetsu Magnetic Observatory (MMB) $\sim$10 km from the Memanbetsu substation. (Bottom) One-second-cadence GIC measurements measured in a grounded neutral point of a Y-connected three-phase transformer connected to a 187 kV bus at the Memanbetsu substation of the Hokkaido Power Co. The subscripts $x$ and $y$ correspond to Geographic North and East, respectively.}
\label{sample}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_model_predictions-MeanModel-20060806T150000.pdf}
\caption{Out-of-sample model predictions (top) and predictions errors for a 1-day interval. Note that the scales for Models 2-4 differ from that for Model 1 so that the error amplitudes for Models 2-4 appear larger than that of Model 1 by a factor of 2.}
\label{predictions}
\end{figure}

@@

The various tests that were performed to determine the influence of each method are summarized below and the associated results are shown in Figures~\ref{a_variations}-~\ref{b_variations}.

@@

\item Five methods for averaging were used to compute magnitudes at each evaluation frequency of a complex-valued model parameter $p(D)$, where $D$ is estimate made using a 1-day of data. If $u(D)$ and $v(D)$ are the real and imaginary phase values for Models 2-4, then the $|p|$ was calculated as: (1) $|\text{mean}(u)+i\text{mean}(v)|$ (the method used), (2) $\text{mean}(|p|)$, (3) $|\text{median}(u)+i\text{median}(v)|$, (4) $\text{median}(|p|)$, and (5) $\text{mloc}(u) + i\text{mloc}(v)$, where the function $\text{mloc}$ is the Huber M-estimator \citep{Huber2011} with $\psi(x)=x$ for $|x|<1.5$ and $\psi(x) = 1.5$ for $x>1.5$ and $\psi(x) = -1.5$ for $x<-1.5$ (the \texttt{mlochuber} function in LIBRA \citep{Verboven2010} was used for the calculation). The out-of-sample PEs varied by $\pm 0.01$ and MSE ratios shown in Table~1 differed by at most $0.03$. Neither averaging method was yielded consistently higher or lower values. Figures~\ref{a_variations}-~\ref{b_variations} shows that most of the sensitivity is below $T=20$ s, as expected, due to the signal-to-noise being near equal to 1 below this period.

@@

\begin{figure}[h]
\begin{minipage}[t]{.48\linewidth}
\includegraphics[width=\textwidth]{figures/plot_options_comparison_a.pdf}
\label{plot_options_comparison_a}
\end{minipage}
\begin{minipage}[t]{.48\linewidth}
\includegraphics[width=\textwidth]{figures/plot_options_comparison_b.pdf}
\end{minipage}
\begin{minipage}[t]{.48\linewidth}
\includegraphics[width=\textwidth]{figures/plot_options_comparison_zx.pdf}
\label{plot_options_comparison_a}
\end{minipage}
\begin{minipage}[t]{.48\linewidth}
\includegraphics[width=\textwidth]{figures/plot_options_comparison_zy.pdf}
\end{minipage}

\end{figure}

>>>>

\begin{abstract}
A common assumption used in estimating geomagnetically induced currents (GICs) in a power system given a nearby direct measurement or indirect estimate (via a transfer function with input of the geomagnetic field) of the horizontal geoelectric field components $E_x$ and $E_y$ on Earth's surface is that the system is resistive. That is, the approximation $GIC(t)$  $=$ $a_oE_x(t) + b_oE_y(t)$ (Model 1) is used, where $a_o$ and $b_o$ are frequency independent network parameters.  We provide the first test of this assumption using GIC measurements made in a 187 kV transformer connected to a $\sim$100~km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$9~km away.  We compute frequency dependent network parameters $a(\omega)$ and $b(\omega)$ in Model 2, $GIC(\omega)$ $=$ $a(\omega)E_x(\omega) + b(\omega)E_y(\omega)$, and find that they are frequency dependent and that this model provides a significantly better representation of measured GIC.
\end{abstract}

@@

Geomagnetically induced currents (GICs) are electric currents in conducting systems due to electric fields induced near Earth's surface. These fields are a result of time variations in ionospheric currents on timescales on the order of hours \citep{Ohtani2000} and the movement of Earth's surface relative to slow-varying current systems in Earth's ionosphere that are near stationary relative to the Sun \citep{Stening2013}. Of particular interest are currents induced in electric power systems as they can lead to system degradation, disruption, and failure \citep{Albertson1993,NERC2012}. Accurate estimation of GICs given direct measurements of the electric field (or estimates of the electric field based on the geomagnetic field) is important for power system design, retrospective analysis, and mitigation of the impacts of space weather on power systems \citep{Molinski2002,Thomson2010,NERC2012,Gaunt2014}.

@@

%\cite{Oyedokun2013a} developed a method to account for the finite inductive response time of transformers by taking the GIC predicted based on the quasi-DC assumption and then applying a piecewise-linear L/R filter derived from bench-scale transformer measurements, with a time constant dependent on load, to obtain an alternative estimate of GIC.

%\todo[inline]{What can I conclude from \cite{Oyedokun2013a}? Mostly the thesis documents visual differences in time series traces with and without accounting for transformers.}

In this work, we use a unique dataset in which measurements $\mathbf{E}$ and the surface horizontal geomagnetic field, $\mathbf{B}$, were made near a site where GICs were being measured in an electric power system.  We first estimate the system parameters $a_o$ and $b_o$ from data using conventional methods, which assume that they are frequency independent. Then, we use a model in which the system parameters are frequency dependent, compute the frequency dependence, and compare this model with the traditional frequency independent model.

@@

The only modification made to the magnetic field measurements was the replacement of 36 data points in all components of the magnetic field on 2006-08-06 with linearly interpolated values. The 36 values for $B_x$ and $B_y$ were the same and much larger than the surrounding values.

@@

\noindent
the matrix equation $\mathbf{GIC} = \underline{\mathbf{E}}\cdot\mathbf{p}$ is created, where $\mathbf{GIC}$ is a 864000x1 matrix of 1 day of observations with zero mean, $\mathbf{p} = [a_o,b_o]^T$, and $\underline{\mathbf{E}}$ is a 86400x2 matrix with rows of $[E_x(t), E_y(t)]$ observations with each column having zero mean. The values of $\mathbf{p}$ are then obtained using a standard linear regression solver. The values of $a_o$ and $b_o$ determined in this way correspond to those that minimize the sum-of-squares of $GIC(t)-G_o(t)$ over a 1-day interval. \citep[][provided the mathematically equivalent closed-form equations.]{Pulkkinen2007}. The averages $\overline{a}_o$ and $\overline{b}_o$ are the average of the 35 $a_o$ and $b_o$ values computed using the 35 unique 1-day window segments of available data.

In Model 2, the system parameters $a(\omega)$ and $b(\omega)$ are frequency dependent according to

@@

The same approach is used for solving for $z_x$ and $z_y$ in Equation~\ref{model4}. For the transfer function $\boldsymbol{Z}$ in $\mathbf{E}(\omega) = \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$ used in Model~3, the $xx$ and $xy$ components of $\boldsymbol{Z}$ are solved for using $E_x(\omega) = Z_{xx}(\omega)B_x(\omega) + Z_{xy}(\omega)B_{y}(\omega)$; the $yx$ and $yy$ components are solved for using $E_y(\omega) = Z_{yx}(\omega)B_x(\omega) + Z_{yy}(\omega)B_y(\omega)$.

@@

There are many statistical methods and variations thereof that have been developed for estimating transfer functions in the magnetotelluric (MT) literature that use additional steps and processing \citep{Egbert1986,Chave1987,Chave1989,Jones1989,Larsen1996,Egbert1997,Eisel2001,Chave2004,Chave2012,Chave2017} beyond the ones used that follow the basic method in \cite{Sims1971} and \cite{Simpson2005}. The objective of these works is typically the estimation of a robust and low bias transfer function. In the results presented, we did not use a robust statistical method or attempt to correct for bias in estimates as our primary interest is comparing the ability of different models to predict the data (in a mean-squared-error sense) rather than estimating the model parameters. As discussed and evaluated in~\cite{Weigel2017}, a model developed using robust and bias correction methods will not necessarily be an optimal prediction model (when the sum-of-squares error is used as a prediction metric).

@@

\begin{table}
\caption{Out-of-sample model performance metrics for models $m=1-4$. MSE$_m$/MSE$_1$ is the ratio of the MSE for model $m$ to that of Model 1. The PE and MSE ratios are the average of 35 values and the confidence interval, CI, was determined from 1000 bootstrap samples of these 35 values.}
\centering
\begin{tabular}{l c c c c}
\hline
$m$\hspace{1em} Model & PE & 95\% CI & MSE$_m$/MSE$_1$ & 95\% CI\\
\hline
1\hspace{1em} $G_o(t) = a_oE_x(t) + b_oE_y(t)$ & 0.35 & [0.25, 0.45] & 1.0 & \\
2\hspace{1em} $G_E(t) = a(\omega)E_x(t) + b(\omega)E_y(t)$ & 0.60 & [0.54, 0.65] & 1.8 & [1.6, 2.0]\\
3\hspace{1em} $G_{E'}(t) = a(\omega)E'_x(t) + b(\omega)E'_y(t)$ & 0.78 & [0.76, 0.80] & 3.1 & [2.8, 3.4]\\
4\hspace{1em} $G_{B}(t) = Z_xB_x(t) + Z_yB_y(t)$ & 0.83 & [0.80, 0.85] & 4.6 & [3.8, 5.3]\\
\hline
\end{tabular}
\end{table}

Figure~\ref{predictions} shows a 1-day interval for selected because the value of the prediction efficiency for each model was in the same order as that shown in Table~1, i.e., Model 1 has the lowest PE and Model 4 has the lowest.

The primary conclusions of this work follow from the results shown in Table~1: the frequency dependent model, Model 2, is significantly better at predicting GIC than the frequency independent model, Model 1. Model 3, which uses an estimate of the electric field based on horizontal magnetic field measurements provides better estimates of GIC than Model 2. Model 4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides the best predictions of GIC.

Although inferior, the Model~1 does have a good prediction performance - the PE is 0.35 and the data/model correlation coefficient of 0.64. The top panel of  Figure~\ref{predictions} shows that its predicted GIC generally follows the overall trend of the measured GIC.

%In a similar way that $|\mathbf{B}|(t)$ or $d|\mathbf{B}|(t)/dt$ are often good predictors of $GIC(t)$ \todo{Pierre: Any suggestions?} {\color{red} \textbf{[REFS]}} despite the fact that physically $\mathbf{E}$ drives GIC and $\mathbf{E}$ is related to $\mathbf{B}$ through a frequency dependent transfer function, Model 1 is a fair predictor of GICs, but is a poor approximation relative to Models 2 and 3.

@@

Consistent with the results in Table~1, The SN ratio for Model 1 is lower than that for Models 2-4 at all frequencies.

Below $600$~s (10~min), Model 2 has the largest SN ratio. This result is somewhat visible in the sample interval shown in Figure~\ref{predictions} where the high frequency fluctuations in the prediction error for Model 2 are smaller than that for Models 3 and 4.

Below $600$~s (10~min), the predictive ability of each model is in the ordering that may be expected based on the data used to derive the model: Model 2, which uses direct measurements of the electric field, has a higher SN ratio than Model 3, which uses an indirect estimate of the electric field. Model 4 has a slightly higher SN than Model 3, and this is expected because they both use the same input of the magnetic field, but Model 3 has an additional source of uncertainty from the parameters in $\mathbf{Z}$. Above $600$~s, the ordering of the SN is not consistent with this expectation and is possibly explained by two factors (1) the measured GIC is based on an integrated electric field and the electric field used in Model 2 is a point measurement - the result is that implicit in the model is an assumption that the electric field is uniform over the length of the line on which GIC is measured and (2) the error associated with this assumption has less of an influence on the errors in a model that uses the point measurements of the magnetic field.

Above $10,000$~s (2.8~hrs), the SN ratio ordering is the same as the ordering of the PE and MSE ratios shown in Table~1, with Model~1 having the lowest and Model~4 having the highest SN ratio.

The average parameters and 95\% CI for Model 1 are $a_o = $ $73 [62,64]$ A/(V/km) and $b_o =$ $-72.0 \pm [-98,-45]$ A/(V/km) and the distribution of the 35 parameters calculated using each 1-day interval is shown in Figure~\ref{histogram}. (Results are similar if 1-minute averages are used; \cite{Watari2015} found $a_o=38.1$ A/(V/km) and $b_o=-7.4$ A/(V/km) for the interval 2006-12-14--2006-12-15 JST time interval).

@@

\section{Discussion}

There are at least three possible explanations of the frequency dependence of the data-derived system parameters: (1) the quasi-DC approximation is invalid, (2) invalid model assumptions, and (3) artifacts of the measurement process.

(1) When the quasi-DC approximation is used, ~\cite{Albertson1981} is usually cited; the justification for the quasi-DC approximation is given the Appendix, where it is stated that ``An analysis based on Carson's equations reveals that transmission lines can be modeled for GIC determination by using their positive sequence resistance values with a small correction factor to account for skin and proximity effects. The correction factor varies from 0.95 to 1.0 depending on conductor size.'' \cite{Lehtinen1985} notes justifies the use of the quasi-DC approximation because GIC amplitudes are significant at frequencies that are much lower than the 50~Hz power line frequency. Carson's equations are equations for the frequency-dependent impedance associated with a wire above a conducting plane. The justification of \cite{Lehtinen1985} is essentially a statement that the frequency dependent component of Carson's impedance is proportional to frequency, the frequency dependent impedance of a power line is small at 50~Hz and so it should be even smaller at GIC frequencies of < 1~Hz. \todo{Need a reference here to 50~Hz values} Thus it appears that the quasi-DC approximation is justified.

(2) The electric field was measured at a single location $\sim$9 km away from the GIC instrument. The exact calculation of GICs in a power system under the quasi-DC assumption requires knowledge of the earth potential at the grounding points of the system, and simulations of GIC flow \citep{Albertson1981,Lehtinen1985} typically start with an assumed electric field along the line to calculate these potential differences. The potential differences are then used along with the system component resistances and connectivity to calculate GIC in the different parts of the system. In both Model~1 and Model~2, the input electric field is assumed to be spatially homogeneous along the line in which GIC was measured. If it is not, the parameters derived from the data in Model~1 may vary significantly from that expected. The same should be true of Model~2, but this modeling error is not expected to introduce a frequency dependence.

(3) A final consideration is that artifacts of the electric field and GIC measurements used in Model~2 cause the observed frequency dependence. Given the response characteristics of these instruments are essentially frequency independent above a period of tens of seconds, this does not explain the frequency dependence observed over the range of 10-10,000~s.

\section{Conclusions}

The primary implications of the results presented are (1) Significantly improved predictions of GICs can be made if frequency dependent system parameters are used and (2) A direct comparison of $a_o$ and $b_o$ estimated using Equation~\ref{model1} and computed using power system configuration information is problematic because the data are not consistent with the assumptions of Equation~\ref{model1}.

Although the use of $a_o$ and $b_o$ values estimated from the data give a reasonable out-of-sample prediction model, it is not known how much these data-estimated values will differ from those directly calculated. (Because the details of the power system under consideration in this work are not available, we were not able make this comparison.)

@@

@@

%Ideally, frequency dependent system parameters could be calculated analytically for comparison with those estimated using data in this work. Because the details of the power system under consideration in this work are not available, we were not able make this comparison. Note that frequency independent system parameters calculated using the approach in this work may not be directly comparable to those found from a network analysis for several reasons: (1) At periods less than $\sim 30$~s, the signal-to-noise ratio is low and the parameter estimates are expected to be biased; as shown in the Appendix, the transfer function estimates are highly sensitive to the method used for their calculation at (2) Also at periods less than $\sim 30$~s, effects of the instrument response may be significant. \cite{Fujii2015} has noted that the 1-s geoelectric field data from the Memambetsu Observatory were low-pass filtered with a 3.3-s cutoff-period prior to 2010; (3) The GIC observed in a power system is based on the integration of the electric field along a power line. All of the models assume that the electric field is constant along the line where the GIC is measured, but many works have shown that the electric field is expected to have variations on length scales comparable to the line considered ($\sim$100~km). \todo{REFS}

\acknowledgments
Magnetic and electric field data made at the Kakioka Magnetic Observatory was obtained from the Japan Meteorological Agency data portal \url{http://www.kakioka-jma.go.jp/obsdata/metadata/en}. We would like to thank the Hokkaido Electric Power Co. for the GIC measurements at the Memanbetsu substation and Shinichi Watari for communicating and preparing the data. 
%\clearpage

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_raw_All_20060805.pdf}
\caption{(Top) One-second-cadence geomagnetic field measured at the Memanbetsu Magnetic Observatory (MMB). (Middle) One-second-cadence geoelectric field measured at the Memanbetsu Magnetic Observatory (MMB) $\sim$9 km from the Memanbetsu substation. (Bottom) One-second-cadence GIC measurements measured in a grounded neutral point of a Y-connected three-phase transformer connected to a 187 kV bus at the Memanbetsu substation of the Hokkaido Power Co. The subscripts $x$ and $y$ correspond to Geographic North and East, respectively.}
\label{sample}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_model_predictions-MeanModel-2006-08-06.pdf}
\caption{Out-of-sample model predictions (top) and predictions errors for a 1-day interval. Note that the scales for Models~2-4 differ from that for Model 1 so that the error amplitudes for Models~2-4 are scaled by a factor of 2 relative to that for Model~1.}
\label{predictions}
\end{figure}

@@

The various tests that were performed to determine the influence of each method are summarized below and the associated results are shown in Figure~\ref{model_2_and_4_variations}.

@@

\item Five methods for averaging were used to compute magnitudes at each evaluation frequency of a complex-valued model parameter $p(D)$, where $D$ is estimate made using a 1-day of data. If $u(D)$ and $v(D)$ are the real and imaginary phase values for Models 2-4, then the $|p|$ was calculated as: (1) $|\text{mean}(u)+i\text{mean}(v)|$ (the method used), (2) $\text{mean}(|p|)$, (3) $|\text{median}(u)+i\text{median}(v)|$, (4) $\text{median}(|p|)$, and (5) $\text{mloc}(u) + i\text{mloc}(v)$, where the function $\text{mloc}$ is the Huber M-estimator \citep{Huber2011} with $\psi(x)=x$ for $|x|<1.5$ and $\psi(x) = 1.5$ for $x>1.5$ and $\psi(x) = -1.5$ for $x<-1.5$ (the \texttt{mlochuber} function in LIBRA \citep{Verboven2010} was used for the calculation). The out-of-sample PEs varied by $\pm 0.01$ and MSE ratios shown in Table~1 differed by at most $0.03$. Neither averaging method was yielded consistently higher or lower values. Figure~\ref{model_2_and_4_variations} shows that most of the sensitivity is below $T=20$ s, as expected, due to the signal-to-noise being near equal to 1 below this period.

@@

\begin{figure}[h]
\begin{minipage}[t]{.48\linewidth}
\includegraphics[width=\textwidth]{figures/plot_options_comparison_a.pdf}
\label{plot_options_comparison_a}
\end{minipage}
\begin{minipage}[t]{.48\linewidth}
\includegraphics[width=\textwidth]{figures/plot_options_comparison_b.pdf}
\end{minipage}
\begin{minipage}[t]{.48\linewidth}
\includegraphics[width=\textwidth]{figures/plot_options_comparison_zx.pdf}
\label{plot_options_comparison_a}
\end{minipage}
\begin{minipage}[t]{.48\linewidth}
\includegraphics[width=\textwidth]{figures/plot_options_comparison_zy.pdf}
\end{minipage}
\caption{Caption}
\label{model_2_and_4_variations}
\end{figure}
